Two-Dimensional Vesicle Hydrodynamics from Hydrophobic Attraction Potential

### 2022-02-15 doc/reviews/ResponseToReviewer3.tex

This is evidence that the lubrication hydrodynamics of the thin film between two particles under the repulsive Morse potential is well captured  if the flow velocity can be accurately calculated in the boundary integral simulations when two particles are sufficiently close to each other and the short-range repulsion is the dominant force.

In this work we used similar numerical techniques to solve the integral equations when some boundaries are quite close to each other.  As the repulsive Morse potential works well to prevent overlapping of particles (provided the time step size is sufficiently small), our previous results in Quaife, Veerapaneni and Young, 2019, Phys. Rev. Fluids gives us confidence in the accuracy of the fluid flow in the thin space between the particles.

\item Change: We added sentences to describe how the lubrication hydrodynamics in a narrow space between boundaries is well-captured within the boundary integral formulation in page 4, starting in line 129.
  Then, in section 4.1 \textit{Model Parameters}, we show quantitatively that the lubrication and
  artificial repulsion are comparable for our simulation parameters.
\end{itemize}

\noindent
\comment{{\bf 2.} Fig.~3 plots the time evolution of the reduced area in (a) and
the excess length of the bilayer structure in (b). Both show obvious
oscillations with multiple frequencies, which are not as smooth as the
inset of (c) showing one dominant frequency. The authors are suggested
to explain the source of multiple frequencies generated in the time
evolution of $A$ and $\Delta$.}
\begin{itemize}
\item Response: We thank the Reviewer for pointing out this feature of
  the curves.  We analyzed the outputs more carefully and concluded that
  the noted frequencies correspond to the tank-treading of the slightly
  polygonal vesicle and particles, respectively.

\item Change: We modified the second to last paragraph in Section 4.1.
  Moreover, we scaled the abscissas of Figure 3 to percent, and corrected
  a scaling factor in the time axes.

### 2022-02-15 doc/main_jfm.tex
main, refs, Review1,3

The second part of the imposed forces and torques comes from repulsion
between proximal particles. Given a pair particles indexed with $i$ and
$j$, we find the two points $\tilde{\xx}_i \in \Gamma_i$ and
$\tilde{\xx}_j \in \Gamma_j$ that are closest to one another (Figure \ref{fig:figure0}a). We then
define the repulsion force and torque
\begin{align}
  \label{eq:REPULforce}
  \FF_i^{\text{repul}} &= \sum_{j \neq i} 
    \frac{\tilde{\xx}_i - \tilde{\xx}_j}
    {|\tilde{\xx}_i - \tilde{\xx}_j|} 
    P'(|\tilde{\xx}_i - \tilde{\xx}_j|), \\
  \label{eq:REPULtorque}
  T_i^{\text{repul}} &= \sum_{j \neq i} 
    (\tilde{\xx}_i - \aa_i)^{\perp} \cdot 
    \frac{\tilde{\xx}_i - \tilde{\xx}_j}
    {|\tilde{\xx}_i - \tilde{\xx}_j|} 
    P'(|\tilde{\xx}_i - \tilde{\xx}_j|).
\end{align}
The repulsion profile $P(r)$ is set to zero for distances $r$ larger than a
repulsion length scale $\rho_0$. As such,~\eqref{eq:REPULforce}
and~\eqref{eq:REPULtorque} ignore particles outside a
$\rho_0$-tubular neighborhood of $\Gamma_i$. For $0 \leq r < \rho_0,$ we use
$P(r) = M(1 - \sin(r/\rho_0))$ where $M$ is sufficiently large to
prevent particle collisions. 
\textcolor{magenta}{
The high-order interpolation-based quadrature rule used to compute the nearly-singular integrals has
been shown to well capture the lubrication hydrodynamics in the narrow space between boundaries 
when the repulsive potential $P$ gives a large steric force \cite{qua-bir2014}.  
In the simulations we adjust the spatial resolution on the particle boundary to ensure that the lubrication hydrodynamics is well resolved as in \cite{qua-vee-you2019,qua-gan-you2021}.
}
%
Then, the total imposed force and torque
are
\begin{equation}
\label{eq:total_forces}
  \FF_i = \FF_i^{\text{hydro}} + \FF_i^{\text{repul}},\quad
  T_i = T_i^{\text{hydro}} + T_i^{\text{repul}}, \qquad
  i=1,\ldots,N_b.
\end{equation}

@@

We have the following jump relations for \eqref{eq:uidef}: 
\begin{equation}
\label{eq:DLjump}
\jump{\ReviewerTwo{\eta}_i} = \sigma, \quad
\jump{w_i} = 0, \quad
\jump{\nabla \ReviewerTwo{\eta}_i} = \frac{\dif \sigma}{\dif s}\ttau,\quad
\jump{\nabla w_i} = \mathbf{0},
\end{equation}
on $\Gamma_i$ (see, e.g. \cite{KlBaGrON13}). Therefore,
\begin{align*}
  \jump{{\bf T}_2 \cdot \nnu}   &= \jump{\SS(\ReviewerTwo{\eta}_i,w_i) \cdot \nnu +
    \SS(w_i,\ReviewerTwo{\eta}_i) \cdot \nnu} \\
  &= \Big[\!\Big[ \left( 2\gamma\rho^{-1} \ReviewerTwo{\eta}_i w_i \mathbf{I} + 2\gamma\rho \nabla \ReviewerTwo{\eta}_i \cdot \nabla w_i \mathbf{I} 
- 2\gamma\rho \nabla \ReviewerTwo{\eta}_i  \nabla w_i^T - 2\gamma\rho \nabla w_i \nabla
  \ReviewerTwo{\eta}_i^T\right) \cdot \nnu \Big]\!\Big] \\
%  &= \jump{\left( 2\gamma\rho^{-1} u_i v_i \mathbf{I} + 2\gamma\rho \nabla u_i \cdot \nabla v_i \mathbf{I} 
%- 2\gamma\rho \nabla u_i  \nabla v_i^T - 2\gamma\rho \nabla v_i \nabla
%  u_i^T\right) \cdot \nnu}\\
&= 2\gamma\rho^{-1} \sigma w_i \nnu + 2\gamma\rho \frac{d\sigma }{ds}\frac{dw_i }{ds} \nnu
- 2\gamma\rho \ \frac{d\sigma }{ds} \frac{dw_i }{d\nnu} \ttau = \JJ_i.
\end{align*}
%\todo[inline]{RR: Can you make sure the last line is correct? There was
%an unbolded $\nu$ which I changed to a bolded $\nnu$.} RJR -> Correct. Thank you.
Combining this with~\eqref{eq:prejump} gives~\eqref{eq:recipforcetorque}
as required.

@@

The base-case is a single JP vesicle consisting of $N_b = 58$ particles
with $c = 0.5 l_0$ for particle radius, $\rho = 2 l_0$ for decay length, $\rho_0 = 0.2l_0$ for repulsion
length, $M=4.0$~$k_BT$ for repulsion strength, $\gamma = \text{4.1 pN
nm}^{-1}$ for interfacial tension.  The parameter $\mu = \text{1 cP} = \text{1 pN
ns nm}^{-2}$ is viscosity of room-temperature water.

@@

\begin{figure}
\begin{center}
  \includegraphics[width=11cm]{Figure5_Wrapper.pdf}\\
  \includegraphics[width=11cm]{Figure5B_Wrapper.pdf}
%\includegraphics[width=0.5\textwidth]{Slip.eps}\\
%\includegraphics[width=0.4\textwidth]{Slip2.eps}
\end{center} 
\caption{\label{figure5} Inter-monolayer slip is present in the JP
  vesicle tank-treading motion at all shear rates. Panel (a) tracks a
  pair of particles (blue and yellow). The shear rate is $\chi=0.005$
  and in the right figure, the yellow particle will complete two and a
  quarter revolutions in the same time the blue particle completes two
  revolutions. The curves in panel (b) are the distance the outer
  leaflet has slid past the inner leaflet. The slopes of the curves
  give the slip velocity. With the exception of $\chi = 0.0035$, the
  slip velocity is generally monotonic in the shear rate.}
\end{figure}

@@

Since the starting configuration is nearly circular, the reduced area
decreases from an initial value close to $1$ and tends to a steady-state
value $A^*_{\infty}$. The data in panel (a) are fit to the model $A^* =
(A^*_{\infty}-a_1-a_2) + a_1 \exp(-k_1t) + a_2 \exp(-k_2t),$ $0 < k_2 <
k_1$. Panel (d) shows that $A^*_{\infty}$ decreases with the shear rate.
The JP vesicle achieves a steady-state reduced area earlier when the
shear rate is low (inset), but the decay rate $k_2$ is more or less
constant for higher shearer rates.

The oscillations in the data of Figure~\ref{figure4} are due to the
granularity of the JP vesicle. The inset of Figure~\ref{figure4}c zooms
in on the arc length data for the shear rate $\chi = 0.003$. It shows
that the oscillations are smooth and well-resolved by our second-order
Adams-Bashforth scheme.

@@

%Table~\ref{table1} provides $b = 0.79 \pm 0.3$ pN ns nm$^{-3}$  
\ReviewerOne{Table~\ref{table1} provides a series of numerical friction coefficients $b$ and the default number of particles is $N_b = 58$. With the fixed total length of the JP vesicle, we change parameters related to the length scale, for instances, $\rho$, $\rho_0$, and the radius of particles.
With $-5\%$, $5\%$, and $10\%$ adjustments in these parameters, the resulting numbers of particles are $N_b=\{61,56,53\}$, respectively. }
%
\ReviewerOne{The overall trend shown in Table~\ref{table1} is values of $b$ decreases as $\rho$
decreases and $b$ are}
%over a range of shear rates which is 
in good quantitative agreement with values previously reported in the literature.
Atomistic studies have also considered inter-monolayer slip in lipid bilayers. 
\cite{WuoEd06} and \cite{denOtter2007} reported $b = 0.7 \times 10^6$ Pa m$^{-1}$ s  
$=0.7$ pN ns nm$^{-3}$  
and $b = 2.4$  pN ns nm$^{-3}$ for DPPC membranes simulated by MD, respectively. 
Using a more recent version of the Martini force field, \cite{Zgorski2019} 
gives $b = 5.5$ pN ns nm$^{-3}$ for shear rates 0.4 ns$^{-1}$ and higher.
It is understandable that there is uncertainty in the friction coefficients of
Table~\ref{table1}.  The scatter in our data, however, is fully consistent with that calculated
from MD simulations, c.f. the transient rise in values of Table~\ref{table1} and in \cite{Zgorski2019}, Figure 10
for low shear rates. 
\subsubsection{Membrane Ruptures}
\begin{figure}
\centering
\includegraphics[width=11.32cm]{Figure8_Wrapper.pdf}
%\includegraphics[height=2in]{N58_rupt_0.pdf}
%\includegraphics[height=2in]{N58_rupt_200.pdf}
\\
%\includegraphics[height=2in]{N58_rupt_400.pdf}
%\includegraphics[height=2in]{N58_rupt_600.pdf}
  \caption{\label{figure8} Tank-treading leads to membrane rupture for
  large shear rates.} 
\end{figure}
A temporary fissure or a complete membrane rupture can occur at large
shear rates. Figure~\ref{figure8} demonstrates how a vesicle can rupture
when suspended in a shear flow. For $\chi = 0.1$, starting with a
circular shape (Figure~\ref{figure8}a), the vesicle is stretched by the
background flow and fissures appear in the bilayer structure in multiple
locations (Figure~\ref{figure8}b). In Figure~\ref{figure8}c, the
ruptured vesicles form two planar micelles which are eventually carried
off by the flow (Figure~\ref{figure8}d). 
%
\ReviewerOne{An emperical numerical study has shown that the critical shear rate for a complete membrane rupture is $\chi=0.0655$. Supplementary Materials Movie 4 shows the membrane rupture at the critical shear rate.}

@@

%The single JP vesicle simulations end with obtaining JP bilayer stretching 
%modulus $K_A$ and hydraulic permeability constant $P$. We extract these 
%physical quantities from the Laplace pressure relationship that $R\Delta P$ is 
%proportional to JP bilayer stretch. We describe the aqueous flux in membrane 
%using a differential equation~\ref{eq:perm} and solve for the theoretical 
%permeability constant $P$. In spite of producing these values with some discrepancy
%using the default parameter set, we provide a new framework in measuring coarse-grained 
%membrane permeability.

>>>>

The second part of the imposed forces and torques comes from repulsion
between proximal particles. Given a pair particles indexed with $i$ and
$j$, we find the two points $\tilde{\xx}_i \in \Gamma_i$ and
$\tilde{\xx}_j \in \Gamma_j$ that are closest to one another (Figure \ref{fig:figure0}a). We then
define the repulsion force and torque
\begin{align}
  \label{eq:REPULforce}
  \FF_i^{\text{repul}} &= \sum_{j \neq i} 
    \frac{\tilde{\xx}_i - \tilde{\xx}_j}
    {|\tilde{\xx}_i - \tilde{\xx}_j|} 
    P'(|\tilde{\xx}_i - \tilde{\xx}_j|), \\
  \label{eq:REPULtorque}
  T_i^{\text{repul}} &= \sum_{j \neq i} 
    (\tilde{\xx}_i - \aa_i)^{\perp} \cdot 
    \frac{\tilde{\xx}_i - \tilde{\xx}_j}
    {|\tilde{\xx}_i - \tilde{\xx}_j|} 
    P'(|\tilde{\xx}_i - \tilde{\xx}_j|).
\end{align}
The repulsion profile $P(r)$ is set to zero for distances $r$ larger than a
repulsion length scale $\rho_0$. As such,~\eqref{eq:REPULforce}
and~\eqref{eq:REPULtorque} ignore particles outside a
$\rho_0$-tubular neighborhood of $\Gamma_i$. For $0 \leq r < \rho_0,$ we use
$P(r) = M(1 - \sin(r/\rho_0))$.
Then, the total imposed force and torque
are
\begin{equation}
\label{eq:total_forces}
  \FF_i = \FF_i^{\text{hydro}} + \FF_i^{\text{repul}},\quad
  T_i = T_i^{\text{hydro}} + T_i^{\text{repul}}, \qquad
  i=1,\ldots,N_b.
\end{equation}

\ReviewerThree{
The high-order interpolation-based quadrature rule used to compute the nearly-singular integrals has
been shown to well capture the lubrication hydrodynamics in the narrow space between boundaries 
when the repulsive potential $P$ gives a large steric force \cite{qua-bir2014}.  
In the simulations we adjust the spatial resolution on the particle boundary to ensure that the lubrication hydrodynamics is well resolved as in \cite{qua-vee-you2019,qua-gan-you2021}.
}
%

@@

We have the following jump relations for \eqref{eq:uidef}: 
\begin{equation}
\label{eq:DLjump}
\jump{\ReviewerTwo{\eta}_i} = \sigma, \quad
\jump{w_i} = 0, \quad
\jump{\nabla \ReviewerTwo{\eta}_i} = \frac{\dif \sigma}{\dif s}\ttau,\quad
\jump{\nabla w_i} = \mathbf{0},
\end{equation}
on $\Gamma_i$ (see, e.g., \cite{KlBaGrON13}). Therefore,
\begin{align*}
  \jump{{\bf T}_2 \cdot \nnu}   &= \jump{\SS(\ReviewerTwo{\eta}_i,w_i) \cdot \nnu +
    \SS(w_i,\ReviewerTwo{\eta}_i) \cdot \nnu} \\
  &= \Big[\!\Big[ \left( 2\gamma\rho^{-1} \ReviewerTwo{\eta}_i w_i \mathbf{I} + 2\gamma\rho \nabla \ReviewerTwo{\eta}_i \cdot \nabla w_i \mathbf{I} 
- 2\gamma\rho \nabla \ReviewerTwo{\eta}_i  \nabla w_i^T - 2\gamma\rho \nabla w_i \nabla
  \ReviewerTwo{\eta}_i^T\right) \cdot \nnu \Big]\!\Big] \\
%  &= \jump{\left( 2\gamma\rho^{-1} u_i v_i \mathbf{I} + 2\gamma\rho \nabla u_i \cdot \nabla v_i \mathbf{I} 
%- 2\gamma\rho \nabla u_i  \nabla v_i^T - 2\gamma\rho \nabla v_i \nabla
%  u_i^T\right) \cdot \nnu}\\
&= 2\gamma\rho^{-1} \sigma w_i \nnu + 2\gamma\rho \frac{d\sigma }{ds}\frac{dw_i }{ds} \nnu
- 2\gamma\rho \ \frac{d\sigma }{ds} \frac{dw_i }{d\nnu} \ttau = \JJ_i.
\end{align*}
%\todo[inline]{RR: Can you make sure the last line is correct? There was
%an unbolded $\nu$ which I changed to a bolded $\nnu$.} RJR -> Correct. Thank you.
Combining this with~\eqref{eq:prejump} gives~\eqref{eq:recipforcetorque}
as required.

@@

The base-case is a single JP vesicle consisting of $N_b = 58$ particles
with $c = 0.5 l_0$ for particle radius, $\rho = 2 l_0$ for decay length,
$\gamma = \text{4.1 pN nm}^{-1}$ for interfacial tension.
The parameter $\mu = \text{1 cP} = \text{1 pN
ns nm}^{-2}$ is viscosity of room-temperature water. 
\ReviewerThree{
  We select \ReviewerOne{$M=4.0$~$\text{pN}$} and $\rho_0 = 0.2l_0$ so that \eqref{eq:REPULforce}
  prevents particle collisions and is comparable to hydrodynamic lubrication.
  \cite{cawthorn_balmforth_2010} provide the asymptotic expression 
  $F_{\text{lubr}} = -\tfrac{3}{2}\pi \mu \dot \epsilon (c/\epsilon)^{3/2}$
  for the lubrication force between two cylinders in the distance of separation $\epsilon \to 0$.
  For our choice of simulation parameters, $F_{\text{lubr}}$ and \eqref{eq:REPULforce} are comparable
  because the particle gaps smaller than the particle radius
  and the relative velocities $\dot \epsilon$ between particles are on
  the order of nm $\text{ns}^{-1}$.
}

@@

\begin{figure}
\begin{center}
  \includegraphics[width=11cm]{Figure5_Wrapper.pdf}\\
  \includegraphics[width=11cm]{Figure5B_Wrapper.pdf}
%\includegraphics[width=0.5\textwidth]{Slip.eps}\\
%\includegraphics[width=0.4\textwidth]{Slip2.eps}
\end{center} 
\caption{\label{figure5} Inter-monolayer slip is present in the JP
  vesicle tank-treading motion at all shear rates. Panel (a) tracks a
  pair of particles (blue and yellow). The shear rate is $\chi=0.005$
  and in the right figure, the yellow particle will complete two and a
  quarter revolutions in the same time the blue particle completes two
  revolutions. The curves in panel (b) are the distance the outer
  leaflet has slid past the inner leaflet. The slopes of the curves
  give the slip velocity. 
  Slip velocity is generally monotonic in the shear rate.
  The case $\chi = 0.003$ is an exception.}
\end{figure}

@@

Since the starting configuration is nearly circular, the reduced area
decreases from an initial value close to $1$ and tends to a steady-state
value $A^*_{\infty}$. The data in panel (a) are fit to the model $A^* =
A^*_{\infty} + a_1 \exp(-k_1t) + a_2 \exp(-k_2t),$ $0 < k_2 <
k_1$. Panel (d) shows that $A^*_{\infty}$ decreases with the shear rate.
The JP vesicle achieves a steady-state reduced area earlier when the
shear rate is low (inset), but the decay rate $k_2$ is more or less
constant for higher shearer rates.

\ReviewerThree{The time courses contain oscillations due to the
granularity of the JP vesicle.
The curve for $\chi = 0.003$ in Figure~\ref{figure4}a, for example, 
has a low-frequency oscillation with a period around $0.7$ $\upmu$s,
and a high-frequency oscillation with a period four times smaller.
The JP vesicle is somwhat hexagonal (see, e.g., Figure~\ref{figure5}a)
and the low-frequency oscillations corresponds to the tank-treading of a 
polygon, rather than smooth, curve. The higher frequency oscillations
comes from the trank-treading of individual particles, that is
it corresponds to the rate of passage of particles through a fixed-angle
axis about the vesicle centre.} 
The inset of Figure~\ref{figure4}c zooms
in on the arc length data for the shear rate $\chi = 0.003$. It shows
that the curves are smooth and well-resolved by our second-order
Adams-Bashforth scheme.

@@

%Table~\ref{table1} provides $b = 0.79 \pm 0.3$ pN ns nm$^{-3}$  
\ReviewerOne{Table~\ref{table1} provides a series of numerical friction coefficients $b$ and the baseline number of particles is $N_b = 58$. With the fixed total length of the JP vesicle, we change parameters related to the length scale, for instances, $\rho$, $\rho_0$, and the radius of particles.
With $-5\%$, $5\%$, and $10\%$ adjustments in these parameters, the resulting numbers of particles are $N_b=\{61,56,53\}$, respectively. }
%
\ReviewerOne{The overall trend shown in Table~\ref{table1} is values of $b$ decreases as $\rho$
decreases and $b$ are}
%over a range of shear rates which is 
in good quantitative agreement with values previously reported in the literature.
Atomistic studies have also considered inter-monolayer slip in lipid bilayers. 
\cite{WuoEd06} and \cite{denOtter2007} reported $b = 0.7 \times 10^6$ Pa m$^{-1}$ s  
$=0.7$ pN ns nm$^{-3}$  
and $b = 2.4$  pN ns nm$^{-3}$ for DPPC membranes simulated by MD, respectively. 
Using a more recent version of the Martini force field, \cite{Zgorski2019} 
gives $b = 5.5$ pN ns nm$^{-3}$ for shear rates 0.4 ns$^{-1}$ and higher.
It is understandable that there is uncertainty in the friction coefficients of
Table~\ref{table1}.  The scatter in our data, however, is fully consistent with that calculated
from MD simulations, c.f. the transient rise in values of Table~\ref{table1} and in \cite{Zgorski2019}, Figure 10
for low shear rates. 
\subsubsection{Membrane Ruptures}
\begin{figure}
\centering
\includegraphics[width=11.32cm]{Figure8_Wrapper.pdf}
%\includegraphics[height=2in]{N58_rupt_0.pdf}
%\includegraphics[height=2in]{N58_rupt_200.pdf}
\\
%\includegraphics[height=2in]{N58_rupt_400.pdf}
%\includegraphics[height=2in]{N58_rupt_600.pdf}
  \caption{\label{figure8} Tank-treading leads to membrane rupture for
  large shear rates.} 
\end{figure}
A temporary fissure or a complete membrane rupture can occur at large
shear rates. Figure~\ref{figure8} demonstrates how a vesicle can rupture
when suspended in a shear flow. For $\chi = 0.1$, starting with a
circular shape (Figure~\ref{figure8}a), the vesicle is stretched by the
background flow and fissures appear in the bilayer structure in multiple
locations (Figure~\ref{figure8}b). In Figure~\ref{figure8}c, the
ruptured vesicles form two planar micelles which are eventually carried
off by the flow (Figure~\ref{figure8}d). 
%
\ReviewerOne{An emperical numerical study has shown that the critical shear rate for a complete membrane rupture is $\chi=0.0655$. Supplementary Materials Movie 4 shows the membrane rupture at the critical shear rate.}

@@

%The single JP vesicle simulations end with obtaining JP bilayer stretching 
%modulus $K_A$ and hydraulic permeability constant $P$. We extract these 
%physical quantities from the Laplace pressure relationship that $R\Delta P$ is 
%proportional to JP bilayer stretch. We describe the aqueous flux in membrane 
%using a differential equation~\ref{eq:perm} and solve for the theoretical 
%permeability constant $P$. In spite of producing these values with some discrepancy
%using the baseline parameter set, we provide a new framework in measuring coarse-grained 
%membrane permeability.
